---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


***
### I. Parameter estimation
#### I.1. Reading the data

```{r}
chlorsulfuron <- read.table('chlorsulfuron.txt', sep='', header = TRUE)
attach(chlorsulfuron)
```


***
#### I.2.1. Writing the objective functions for the different dose response models

- Four-parameter log-logistic model
- Brain-Cousens hormesis model
- Cedergreen-Ritz-Streibig modified log-logistic model


```{r}
# Four-parameter log-logistic function
Fparfun <- function(xdata, thetas) {
  e <- thetas[1] + (thetas[2] - thetas[1]) / (1 + exp(thetas[3] * (log(xdata) - log(thetas[4]))))
  return(e)
}


# Brain-Cousens hormesis function
Brainfun <- function(xdata, thetas) {
  e <- thetas[1] + (thetas[2] - thetas[1] + thetas[5] * xdata) / (1 + exp(thetas[3] * (log(xdata) - log(thetas[4]))))
  return(e)
}


# Cedergreen-Ritz-Streibig modified log-logistic function
Cederfun <- function(xdata, thetas) {
  e <- thetas[1] + (thetas[2] - thetas[1] + thetas[5] * exp(-1/(xdata**thetas[6]))) / (1 + exp(thetas[3] * (log(xdata) - log(thetas[4]))))
  return(e)
}
```


***
#### I.2.2. Writhing the objective function J
Use residual sum of squares as objective function
```{r}
# objective function
Jtheta <- function(thetas, FUN, xdata, ydata) {
  ydatahat <- do.call(FUN, list(xdata, thetas))
  yresid <- ydata - ydatahat
  J <- sum(yresid^2)
  return(J)
}
```


***
#### I.3. Minimizing the objective function
#### I.3.1 Finding initial values

- C: The horizontal asymptote on the right side (for very large values values of x).
- D: The horizontal asymptote on the left side (for very small values of x).
- b: The steepness - absolute value of slope for a fitted line - of the dose-response curve.
- e: The dose to reach 50% of the maximal eﬀect. **Use initial parameter value e = 1.**
- F: The size of the ‘hormesis’ eﬀect. This is a positive respone at low doses of a toxin or stressor. The larger the value is, the larger the hormesis eﬀect. F = 0 corresponds to no hormesis eﬀect and the resulting model is the four-parameter log-logistic model. **Use initial parameter value F = 1**.
- α: a freely varying parameter. **Use initial parameter value α = 1**.


```{r}
# finding initial b value
m <- lm(log(area) ~ dose)
plot(dose, log(area))
abline(m)
```
```{r}
summary(m)
```
**b** = abs(-0.68547) = **0.6855**

In order to identify initial value of b, fitted line of the dose response curve was drawn and its corresponding slope was labeled. R squared value of the linear function used 
to draw the fitted line is poor mainly due to the non-linearity of the data. 


***
#### I.3.2 Setting initial values
```{r warning=FALSE}
pFpar <- c(min(log(area)), max(log(area)), 0.6855, 1) #(C D, b, e)
pBrain <- c(min(log(area)), max(log(area)), 0.6855, 1, 1) #(C D, b, e, F)
pCeder <- c(min(log(area)),  max(log(area)), 0.6855, 1, 1, 1) #(C D, b, e, F, α)
```

Initial values of the parameter was chosen under following rationale. Minimum of log transformed area value was chosen as initial value of the parameter C, because all models 
converge to log(y) = C at very large value of x or dose. Also, it has been mentioned that C is asymptotic value at very small value of x, so smallest log transformed y value 
was chosen as the value of C. On the other hand, maximum log transformed y value was chosen as initial value of D, because D is the asymptotic value when x is very large. The 
initial values for e, F and α were already set and value of b was taken from fitted line slope. 


***
#### I.3.3 Minimizing the objective function using 'Nelder-Mead algorithm'
For funtion 'optim', **convergence = 0** indicates successful completion.
```{r warning=FALSE}
Fparfit <- optim(par = pFpar, fn = Jtheta, FUN = Fparfun, xdata = dose, ydata = log(area)) # convergence = 0
Fparfit 
```

```{r warning=FALSE}
Brainfit <- optim(par = pBrain, fn = Jtheta, FUN = Brainfun, xdata = dose, ydata = log(area), control = list(maxit=1000)) # convergence = 0
Brainfit
```

```{r warning=FALSE}
Cederfit <- optim(par = pCeder, fn = Jtheta, FUN = Cederfun, xdata = dose, ydata = log(area), control = list(maxit=1000)) # convergence = 0
Cederfit
```

For all models, we have $convergence of zero, representing that all iterations were completed successfully. 




***
### II. Plot the data and fitted curves

Plot the callus area in function of the dose along with the fitted curves
```{r}
plot(x = dose, y = log(area),
     ylim = c(0, 8), xlim = c(0, 10),
     ylab = "area (mm**2)", xlab = "dose (nmol/L)",
     pch = 19)
xseq <- seq(min(dose), max(dose), 0.1)
lines(x = xseq, y = Fparfun(xseq, Fparfit$par), lty = "dashed", col = 'green')
lines(x = xseq, y = Brainfun(xseq, Brainfit$par), lty = "solid", col = 'orange' )
lines(x = xseq, y = Cederfun(xseq, Cederfit$par), lty = "dotdash", col = 'blue')
legend(x = "topright", 
       legend = c("Four-parameter log-logistic function", "Brain-Cousens hormesis function", 
                  "Cedergreen-Ritz-Streibig modified log-logistic function"), 
       lty = c("dashed", "solid", "dotdash"), 
       col = c('green', 'orange', 'blue'))
```




***
### III. Select the best model
Select the best model based on both the Bayesian Information Criterion (BIC) and F-tests. (Significance level of 5%)

#### III.1.1 Functions to calculate Bayesian Information Criterion (BIC)

```{r}
# Bayesian Information Criterion (BIC)
BIC <- function(thetas, FUN, xdata, ydata) {
  RSS <- Jtheta (thetas, FUN, xdata, ydata)
  N <- length(xdata)
  p <- length(thetas)
  B <- N * log(RSS / N) + p * log(N)
  return(B)
}
```


***
#### III.1.2 Function to perform a F-test, statistical hypothesis testing

```{r}
# F-test
Ftest <- function(thetas1, thetas2, FUN1, FUN2, xdata, ydata) {
  
  RSS1 <- Jtheta(thetas1, FUN1, xdata, ydata)
  RSS2 <- Jtheta(thetas2, FUN2, xdata, ydata)
  
  N <- length(xdata)
  p1 <- length(thetas1)
  p2 <- length(thetas2)
  F <- ((RSS1 - RSS2) / (p2 - p1)) / (RSS2 / (N - p2))
  
  pvalue <- pf(F, df1 = p2 - p1, df2 = N - p2, lower.tail = FALSE)
  return(list(F = F, df1 = abs(p2 - p1), df2 = N - p2, pValue = pvalue))
}
```


***
#### III.2 Selection of the model
#### III.2.1 Check BIC values
Select model with lowest value of criterion
```{r}
BIC(Fparfit$par, Fparfun, dose, log(area)) # the lowest
BIC(Brainfit$par, Brainfun, dose, log(area))
BIC(Cederfit$par, Cederfun, dose, log(area))
```

Four-parameter log-logistic model has the lowest BIC value among three. This results represent that among the optimized model generated from set models, four parameter log-
logistic model gave best goodness of fit to given data. The equation that generates BIC value  is designed to give lower (more negative) value to model that gives best fit. 
Now we verify this with the F-test.


***
#### III.2.2 Perform F-tests
Peform F-tests at a significance level of 5%
```
H0: No significant difference between simple and complex model
HA: Complex model is significantly better than the simple model
```
```{r}
Ftest(Fparfit$par, Brainfit$par, Fparfun, Brainfun, dose, log(area))
qf(0.95,1,38)
```
For F-test between:

 - Four-parameter log-logistic model (simple model)
 - Brain-Cousens hormesis model (complex model)

p-value > 5%, null hypothesis cannot be rejected. As shown, F-statistics value from the test was less than critical value (0.6537327 < 4.098172). There is no significant 
difference between simple and complex model. Complex model is not better than simple model.



```{r}
Ftest(Fparfit$par, Cederfit$par, Fparfun, Cederfun, dose, log(area))
qf(0.95,2,37)
```

For F-test between:

 - Four-parameter log-logistic model (simple model)
 - Cedergreen-Ritz-Streibig modified log-logistic model (complex model)

p-value > 5%, null hypothesis cannot be rejected. As shown, F-statistics value from the test was less than critical value (0.062818 < 3.251924). There is no significant difference between simple and complex model. Complex model is not better than simple model.


***
**Conclusion on selection:**

Based on both BIC and F-tests, **Four-parameter log-logistic model** is selected as the best model. Four-parameter log-logistic model has the lowest BIC value and no complex 
model is gives signficantly better fit than this model.




***
### IV. Perform a local sensitivity analysis
Selected model: Four-parameter log-logistic model

```{r}
# define function to calculate sensitivity
sensitivity <- function(FUN, thetas, xdata, eps = 1e-5) {
  yhat <- do.call(FUN, list(xdata, thetas))
  S <- matrix(NA, nrow = length(xdata), ncol = length(thetas))
  for (p in 1:length(thetas)) {
    thetas. <- thetas
    thetas.[p] <- thetas[p] * (1+eps)
    yhat. <- do.call(FUN, list(xdata, thetas.))
    S[, p] <- (yhat. - yhat) / (eps*thetas[p])
  }
  
  return(S)
}


# define function to plot sensitivity
plotSens <- function(sens, parNames = c("C", "D", "b", "e")) {
  require(ggplot2)
  require(reshape2)
  
  Sens <- as.data.frame(sens)
  colnames(Sens) <- parNames
  Sens.m <- melt(data.frame(x = seq(from=0, to=10, by=0.1), Sens), id.vars = "x")
  
  gg <- ggplot(Sens.m, aes(x = x, y = value)) +
    geom_line() + ylab(expression(paste("Sensitivity ",bgroup("(",over(delta~y[j],delta~theta[i]),")")))) + xlab("Dose") +
    facet_grid(variable ~ ., scales = 'free', labeller = label_parsed)+
    theme_bw()+theme(strip.background = element_rect(fill = "white"), strip.text.y = element_text(angle = 0))
  
  return(gg)
}
```

```{r}
sensFpar <- sensitivity(FUN = Fparfun, thetas = Fparfit$par, xdata = seq(from=0, to=10, by=0.1), eps = 1e-5)

plotSens(sens = sensFpar, parNames = c("C", "D", "b", "e"))
```

***
Based on sensitivity plot, each parameter's senstitivy will be first analyzed. Sensitivity of parameter C starts from zero and converges towards 1 as x approaches large 
value. Sensitivity of 1 represents that deviation in parameter value is has direct impact on outcome variable (area) with slope of 1. This represents that initially, change 
in C has no impact on area but its impact on y slowly increases until it plateaus to 1. This result can be explained analytically also, as when x approaches zero, the 
function transforms into y = C + (D - C) / 1 which is the same as y = D. Notice that C disappears from the equation, thus it makes sense that its sensitivity is zero when 
value of x is really small. On the other hand, when x approaches very large value (assumed to be infinite for simplicity sake), the equation transforms into y = C + (D -c) / 
∞, which is y = C. This explains the sort of linear relationship between the value of y and C when x approaches large value. 

For D, it is the exact opposite. Initially, change in value of D has unit impact (1) on value of area, but its impact decreases down to 0, meaning it has no impact, when 
value of x increases. As explained previously, when value of x is really small, the equation transforms into y = D. This explains sensitivity of 1 near x = 0. When value of x 
increases, the equation transforms into y = D, thus also explaining why sensitivity of **zero** is achieved at large values of x. 

For b, we observe three patterns. From x = 0 to x = 0.3, sensitivity increases from 0 to 0.75. The reason why it starts from zero is due to the function transforming into y = 
D at very small value of x. Then sensitivity decreases to negative value when x goes from 0.3 to 2. Negative sensitivity represents inverse relationship between deviation in 
parameter value and outcome variable value. Finally from x = 2 onwards, sensitivity approaches zero. The reason why it approaches zero at very large value of x is due to 
transformation of equation into y = C. 

For e, there were two changes in slope. From x = 0 to x = 1, it's sensitivity increases to over 3. Then it converges to zero. The reason why it starts from zero and end up 
converging to zero was explained above. 

Sensitivity analysis is aimed to identify range of input value where sensitivity is significantly large. Identifying those regions and taking measurements in those regions 
may be necessary to generate better parameter values. In this case, sensitivity of most parameters ranged from -1 to 1, but e spikes up to 3 at x value of around 1. Taking 
more measurements from x = 0.75 to 1.25 may be deemed necessary. 




***
### V. Calculate 95% CI for the parameter estimation of selected model
```{r}
# Define a function to get Fisher Information Matrix (FIM)
FisherInformation <- function(thetas, FUN, xdata, eps = 1e-5) {
  sens <- sensitivity(FUN, thetas, xdata, eps)
  FIM <- t(sens)%*%sens
  return(FIM)
}


# Define a function to get confidence interval (CI)
confInterval <- function(thetas, FUN, xdata, alpha = 0.05, eps = 1e-5) {
  FIM <- FisherInformation(thetas, FUN, xdata, eps)
  C <- solve(FIM) 
  StDev <- sqrt(diag(C))
  N <- length(xdata)
  p <- length(thetas)
  t.05 <- qt(p = 1-alpha/2, df = N-p)
  CI <- data.frame(lower.CI = thetas - StDev*t.05, upper.CI = thetas + StDev*t.05)
  return(CI)
}
```

```{r}
confInterval(thetas = Fparfit$par, FUN = Fparfun, xdata = dose)
```


For each of parameter, a confidence interval can be derived.

The purpose of confidence intervals is to give us a range of values for our estimated population parameter rather than a single value or a point estimate.

Accoding to the result, We are 95% confident that the true value of the parameters are in our confidence interval;

- 95% confident that the true value of **C** is in CI = [-0.1407856, 1.751977]
- 95% confident that the true value of **D** is in CI = [6.9369061,	7.818471]
- 95% confident that the true value of **e** is in CI = [0.7258561,	2.898664]
- 95% confident that the true value of **f** is in CI = [0.5610417,	1.178931]


***
**conclusion:**

95 % CI of parameter values were calculated and their lower and upper bounds are shown. The precision of the estimates were determined by obtaining covariance matrix from the 
inverse Fisher Information Matrix and obtaining the standard deviation off of it. The analysis of the results lead to a conclusion that parameters C and e were within the 
confidence interval(CI) or were not significantly different from the parameter value generated by Four parameter log-logistic function while vice versa for D and b.

***


