---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
## Project2: Linear Regression

```
file.show("data.csv")
```


**Model depned on goals**

The goal is to use multiple linear regression for estimating brain size (measured by ECV) from external skull measurements and body mass.

--> Determine effects of an exposure on an outcome.

**General**

Y = outcome = ECV
X = variables = N.O, Bs, Bcal, Height, Width, AvgBodyMass, (Infraclass)




***
#### (a) Explanatory Analysis on the data

```{r warning=FALSE}
Data <- read.csv('data.csv') # Loading the file
attach(Data)
```


**[1] Filter data depends on Infraclasses**

```{r message=FALSE, warning=FALSE}
library('dplyr')
```

```{r}
Monotremata <- filter(Data, Infraclass == "Monotremata")
head(Monotremata)
```

```{r}
Marsupiala <- filter(Data, Infraclass == "Marsupiala")
head(Marsupiala)
```

```{r}
Placental <- filter(Data, Infraclass == "Placental")
head(Placental)
```



***
**[2] Simple Scatterplots**

Simple scatterplot of ECV & N.O
```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(N.O, ECV)
plot(N.O, log(ECV))
plot(log(N.O), ECV)
plot(log(N.O), log(ECV)) # shows linearity
```

```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(log(N.O), log(ECV), main='All together')
with(Monotremata, plot(log(N.O), log(ECV), main='Monotremata'))
with(Marsupiala ,plot(log(N.O), log(ECV), main='Marsupiala'))
with(Placental, plot(log(N.O), log(ECV), main='Placental'))
```





Simple scatterplot of ECV & Bs
```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(Bs, ECV)
plot(Bs, log(ECV))
plot(log(Bs), ECV)
plot(log(Bs), log(ECV)) # shows linearity
```

```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(log(Bs), log(ECV), main='All together')
with(Monotremata, plot(log(Bs), log(ECV), main='Monotremata'))
with(Marsupiala ,plot(log(Bs), log(ECV), main='Marsupiala'))
with(Placental, plot(log(Bs), log(ECV), main='Placental'))
```





Simple scatterplot of ECV & Bcal
```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(Bcal, ECV)
plot(Bcal, log(ECV))
plot(log(Bcal), ECV)
plot(log(Bcal), log(ECV)) # shows linearity
```

```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(log(Bcal), log(ECV), main='All together')
with(Monotremata, plot(log(Bcal), log(ECV), main='Monotremata'))
with(Marsupiala ,plot(log(Bcal), log(ECV), main='Marsupiala'))
with(Placental, plot(log(Bcal), log(ECV), main='Placental'))
```





Simple scatterplot of ECV & Height
```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(Height, ECV)
plot(Height, log(ECV))
plot(log(Height), ECV)
plot(log(Height), log(ECV)) # shows linearity
```

```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(log(Height), log(ECV), main='All together')
with(Monotremata, plot(log(Height), log(ECV), main='Monotremata'))
with(Marsupiala ,plot(log(Height), log(ECV), main='Marsupiala'))
with(Placental, plot(log(Height), log(ECV), main='Placental'))
```





Simple scatterplot of ECV & Width
```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(Width, ECV)
plot(Width, log(ECV))
plot(log(Width), ECV)
plot(log(Width), log(ECV)) # shows linearity
```

```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(log(Width), log(ECV), main='All together')
with(Monotremata, plot(log(Width), log(ECV), main='Monotremata'))
with(Marsupiala ,plot(log(Width), log(ECV), main='Marsupiala'))
with(Placental, plot(log(Width), log(ECV), main='Placental'))
```





Simple scatterplot of ECV & AvgBodyMass
```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(AvgBodyMass, ECV)
plot(AvgBodyMass, log(ECV))
plot(log(AvgBodyMass), ECV)
plot(log(AvgBodyMass), log(ECV)) # shows linearity
```

```{r warning=FALSE}
par(mfrow=c(2, 2))
plot(log(AvgBodyMass), log(ECV), main='All together')
with(Monotremata, plot(log(AvgBodyMass), log(ECV), main='Monotremata'))
with(Marsupiala ,plot(log(AvgBodyMass), log(ECV), main='Marsupiala'))
with(Placental, plot(log(AvgBodyMass), log(ECV), main='Placental'))
```


***
Drawing scatterplot of predictor variable vs response variable reveals that there is a linear relationship of varying degree between predictor variable and response 
variable. 

Plotting each variable(X) with outcome(Y), the data seems highly skewwed in left. So, for linearity, transformation of variable can be considered.

For all six variables, linearity resumed with taking log-transformation on both variable and outcome: log(Y) ~ log(X)

Hence, we take the **log-transformed variables** for the model building.

***
After the transformation, in general they were positively correlated, meaning that increase in predictor variable led to increase in response variable. 

This makes logical sense as animlas of bigger height, for example, are likely to have bigger skull, thus higher ECV. Overall, there was a clear linearity between predictor 
and response variable with positive slope.  










***
#### (b) Building Regression model

**Foward model building**

- Include predictor one by one
- The included predictor is ideally the one that is most strongly associated with outcome, after verification for the other predictors in the model
- For including, compare absolute t-value between variables of p-value < 5%(~10%), including interaction term.
- Include one by one, first from variable with the highest absolute t-value


```{r message=FALSE, warning=FALSE}
library(car) # load car to use vif function
```


***
#### Step 1: choose first variable according to t-value
Starting from the empty model, **lm(ECV ~ 1)**

Candidates for adding are:

- N.O
- Bs
- Bcal
- Height
- Width
- AvgBodyMass


```{r}
model_11 <- lm(log(ECV) ~ log(N.O))
summary(model_11)
```

```{r}
model_12 <- lm(log(ECV) ~ log(Bs))
summary(model_12)
```

```{r}
model_13 <- lm(log(ECV) ~ log(Bcal))
summary(model_13)
```

```{r}
model_14 <- lm(log(ECV) ~ log(Height))
summary(model_14)
```

```{r}
model_15 <- lm(log(ECV) ~ log(Width))
summary(model_15)
```

```{r}
model_16 <- lm(log(ECV) ~ log(AvgBodyMass))
summary(model_16)
```


***
**t-value**

- N.O. = 52.34, p-value < 5% --> significant
- BS = 29.27, p-value < 5% --> significant
- Bcal = 40.73, p-value < 5% --> significant
- Height = 108.9, p-value < 5% --> significant
- **Width = 109.96**, p-value < 5% --> significant
- AvgBodyMass = 24.48, p-value < 5% --> significant


So, we start building model with Width. Width is added to the empty model. 




***
#### Step 2: choose second variable
Starting from **lm(log(ECV) ~ log(Width))**

Candidates for adding are:

- N.O
- Bs
- Bcal
- Height
- AvgBodyMass


```{r}
model_21 <- lm(log(ECV) ~ log(Width) + log(N.O))
summary(model_21)
```

```{r}
model_22 <- lm(log(ECV) ~ log(Width) + log(Bs))
summary(model_22)
```

```{r}
model_23 <- lm(log(ECV) ~ log(Width) + log(Bcal))
summary(model_23)
```

```{r}
model_24 <- lm(log(ECV) ~ log(Width) + log(Height))
summary(model_24)
```

```{r}
model_25 <- lm(log(ECV) ~ log(Width) + log(AvgBodyMass))
summary(model_25)
```


***
**t-value**

- N.O = 9.586, p-value < 5% --> significant
- Bs = 3.247, p-value = 0.00137 < 5% --> significant
- Bcal = 4.941, p-value = 1.66e-06 < 5% --> significant
- **Height = 13.21**, p-value < 5% --> significant
- AvgBodyMass = 1.673, but p-value = 0.986 > 10% --> we do not consider AvgBodyMass


Comparing absolute t-values, Height has the largest. Hence, Height is added to the model.




***
#### Step 3: choose the third variable
Starting from **lm(log(ECV) ~ log(Width) + log(Height))**

Candidates for adding are:

- N.O
- Bs
- Bcal
- AvgBodyMass


```{r}
model_31 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O))
summary(model_31)
```
- for N.O, t-value is 4.622
- p-value is 6.9e-06, which is <5% --> significant


```{r}
model_32 <- lm(log(ECV) ~ log(Width) + log(Height) + log(Bs))
summary(model_32)
```
- for Bs, t-value is 2.099
- p-value is 0.0371, which is <5% --> significant


```{r}
model_33 <- lm(log(ECV) ~ log(Width) + log(Height) + log(Bcal))
summary(model_33)
```
- for Bcal, t-value is 3.148
- p-value is 0.0019, which is <5% --> significant


```{r}
model_34 <- lm(log(ECV) ~ log(Width) + log(Height) + log(AvgBodyMass))
summary(model_34)
```
- for AvgBodyMass, **p-value is 0.449, which is higher than 10%**
- So we do not consider AvgBodyMass


***
**t-value**

- N.O =  4.622
- Bs = 2.099
- Bcal = 3.148


Comparing absolute t-values, N.O has the higest. Hence, N.O is added to the model.




***
#### Step 4
Starting from **lm(log(ECV) ~ log(Width) + log(Height) + log(N.O))**

Candidates for adding are:

- Bs
- Bcal
- AvgBodyMass


```{r}
model_41 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(Bs))
summary(model_41)
```
- for Bs, **p-value is 0.969, which is higher than 10%**
- So we do not consider Bs


```{r}
model_42 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(Bcal))
summary(model_42)
```
- for Bcal, **p-value is 0.439295, which is higher than 10%**
- So we do not consider Bcal


```{r}
model_43 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass))
summary(model_43)
```
- for AvgBodyMass, t-value is -2067
- p-value is 0.0401, which is <5% --> significant


***
**t-value**

- AvgBodyMass = -2.067

AvgBodyMass is only significant and considerable. So, AvgBodyMass is added to the model.




***
#### Step 5
Starting from **lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass))**

Candidates for adding are:

- Bs
- Bcal


```{r}
model_51 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Bs))
summary(model_51)
```
- for Bs, **p-value is 0.7053, which is higher than 10%**
- so, we do not consider Bs


```{r}
model_52 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Bcal))
summary(model_52)
```
- for Bcal, **p-value is 0.7053, which is higher than 10%**
- so, we do not consider Bcal


Hence, no more variable is significant and considerable. So, from now on, the interaction term between two variables are considered.




***
#### Step 6
Starting from **lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass))**

Candidates for adding are:

- Width:Height
- Width:N.O
- Width:AvgBodyMass
- Height:N.O
- Height:AvgBodyMass
- N.O:AvgBodyMass


```{r}
model_61 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Width):log(Height))
summary(model_61)
```
- for Width:Height, t-value is -4.083
- p-value is 6.5e-05 < 5% --> significant


```{r}
model_62 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Width):log(N.O))
summary(model_62)
```
- for Width:N.O, t-value is -4.302
- p-value is 2.69e-05 < 5% --> significant


```{r}
model_63 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Width):log(AvgBodyMass))
summary(model_63)
```
- for Width:AvgBodyMass, t-value is -4.552
- p-value is 9.39e-06 < 5% --> significant


```{r}
model_64 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(N.O))
summary(model_64)
```
- for Height:N.O, t-value is -4.856
- p-value is 2.47e-06 < 5% --> significant


```{r}
model_65 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass))
summary(model_65)
```
- for Height:AvgBodyMass, t-value is -5.197
- p-value is 5.13e-07 < 5% --> significant


```{r}
model_66 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(N.O):log(AvgBodyMass))
summary(model_66)
```
- for N.O:AvgBodyMass, t-value is -5.091
- p-value is 8.42e-07 < 5% --> significant


**t-value**

- Width:Height = -4.083
- Width:N.O = -4.302
- Width:AvgBodyMass = -4.552
- Height:N.O = -4.856
- **Height:AvgBodyMass = -5.197**
- N.O:AvgBodyMass = -5.091


Comparing absolute t-values, Height:AvgBodyMass has the higest. Hence, Height:AvgBodyMass is added to the model.




***
#### Step 7
Starting from **lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass))**

Candidates for adding are:

- Width:Height
- Width:N.O
- Width:AvgBodyMass
- Height:N.O
- N.O:AvgBodyMass


```{r}
model_71 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(Height))
summary(model_71)
```
- for Width:Height, **p-value is 0.959, which is higher than 10%**
- so, we do not consider Width:Height


```{r}
model_72 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(N.O))
summary(model_72)
```
- for Width:N.O, **p-value is 0.56779, which is higher than 10%**
- so, we do not consider Width:N.O


```{r}
model_73 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))
summary(model_73)
```
- for Width:AvgBodyMass, t-value is 3.284
- p-value is 0.00122 < 5% --> significant


```{r}
model_74 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Height):log(N.O))
summary(model_74)
```
- for Height:N.O, **p-value is 0.6009, which is higher than 10%**
- so, we do not consider Height:N.O


```{r}
model_75 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(N.O):log(AvgBodyMass))
summary(model_75)
```
- for N.O:AvgBodyMass, **p-value is 0.90592, which is higher than 10%**
- so, we do not consider N.O:AvgBodyMass


**t-value**

- **Width:AvgBodyMass = 3.284**


Width:AvgBodyMass is only significant and considerable. So, Width:AvgBodyMass is added to the model.




***
#### Step 8
Starting from **lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))**

Candidates for adding are:

- Width:Height
- Width:N.O
- Height:N.O
- N.O:AvgBodyMass


```{r}
model_81 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Width):log(Height))
summary(model_81)
```
- for Width:Height, **p-value is 0.71624, which is higher than 10%**
- so, we do not consider Width:Height

```{r}
model_82 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Width):log(N.O))
summary(model_82)
```
- for Width:N.O, **p-value is 0.95897, which is higher than 10%**
- so, we do not consider Width:N.O


```{r}
model_83 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Height):log(N.O))
summary(model_83)
```
- for Height:N.O, **p-value is 0.982216, which is higher than 10%**
- so, we do not consider Height:N.O

```{r}
model_84 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(N.O):log(AvgBodyMass))
summary(model_84)
```
- for N.O:AvgBodyMass, **p-value is 0.84109, which is higher than 10%**
- so, we do not consider N.O:AvgBodyMass


Hence, no more variable is significant and considerable. So, from now on, the interaction term between three variables are considered.




***
#### Step 9
Starting from **lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))**

Candidates for adding are:

- Width:Height:N.O
- Width:Height:AvgBodyMass
- Width:N.O:AvgBodyMass
- Height:N.O:AvgBodyMass


```{r}
model_91 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Width):log(Height):log(N.O))
summary(model_91)
```
- for Width:Height:N.O, **p-value is 0.76189, which is higher than 10%**
- so, we do not consider Width:Height:N.O


```{r}
model_92 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Width):log(Height):log(AvgBodyMass))
summary(model_92)
```
- for Width:Height:AvgBodyMass, **p-value is 0.444105, which is higher than 10%**
- so, we do not consider Width:Height:AvgBodyMass


```{r}
model_93 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Width):log(N.O):log(AvgBodyMass))
summary(model_93)
```
- for Width:N.O:AvgBodyMass, **p-value is 0.459936, which is higher than 10%**
- so, we do not consider Width:N.O:AvgBodyMass


```{r}
model_94 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Height):log(N.O):log(AvgBodyMass))
summary(model_94)
```
- for Width:N.O:AvgBodyMass, **p-value is 0.31676, which is higher than 10%**
- so, we do not consider Height:N.O:AvgBodyMass


Hence, no more variable is significant and considerable. So, from now on, the interaction term between four variables are considered.




#### Step 10
Starting from **lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))**

Candidate for adding is:

- Width:Height:N.O:AvgBodyMass


```{r}
model_101 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + log(Width):log(Height):log(N.O):log(AvgBodyMass))
summary(model_101)
```
- for Width:Height:N.O:AvgBodyMass, **p-value is 0.338793, which is higher than 10%**
- so we do not consider Width:Height:N.O:AvgBodyMass



***
Hence, no more variable is significant or considerable. No more variable can be added, so the model building is finalized: 

**lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))**

***
During this analysis, t-value was used as an indicator for forward model building. As the purpose of this model is to provide most accurate prediction of response varialble, 
multicollinearity among the terms were not initially considered. However, the final model's VIF values were all less than 10, hinting that model accurately predicts the 
value of response variable with no major collinear effect.










***
#### (c) Adding variable 'Infraclass'

Final model from regreeesion model building in (b):
```lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))```


Infraclass = categorical variables

- Marsupials
- Monotremes
- Placentals

so, create 2 dummy variable by choosing 1 as reference:

- Marsupials = (0, 0)
- Monotremes = (1, 0)
- Placentals = (0, 1)


```{r}
m <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + factor(Infraclass))
summary(m)
```
Dummy variables are having high p-value, which are higher than 5%, except placenta. This may indicate that model predicts more accurately when phylogenetic similarity among 
placenta is taken into account. However, low p value in other infraclasses indicate that taking phylogenetic similarity is significant in model prediction. The reasons for 
such can be attributed to variance in sample sizes, for example, there are only 2 species that belong to monotremata in this data set. Overall, infraclass seems not 
significant. This can be verified from F-test, following:


--> **Partial F-test**

Two nested models:

- Complete model(mC)
  - with Pc = 9 (intercept + 6 variables + 2 dummy)
- Reduced model(mR)
  - with Pr = 7 (intercept + 6 variables)

```  
H0: There is no significant difference between mC and mR

```

```{r}
mC <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass) + factor(Infraclass))
mR <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))
anova(mC, mR) # F-test
```
- p-value is 0.05188, which is higher than 5%
- According to the F-test, HO cannot rejected
  - There are no significant difference between complete model(mC) and reduced model(mR)

However, due to relative low p-value of the test. It can still be assumed that there is a certain assocation between phylogenetic classes and ECV prediction. 


Hence, the effect of adding categorical variable Infraclass as a predictor is not significant.





***
#### (d) Check the assumptions

**Summary Residual plots:**

  - Residual vs Fitted: Linearity
  - Normal Q-Q: Normality
  - Scale-Location: Homoscedasticity 
  - Residuals vs leverage: Outliers/ Influential points

```{r warning=FALSE}
final_model <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))

plot(final_model, 1)


plot(final_model, 2)


plot(final_model, 3)


plot(final_model, 5)
```


***
Major assumptions of linear regression will be analyzed through analysis of four graphs above. First of all the linearity of the data can be checked by looking at Residual 
vs fitted graph above. If there is a pattern, either positive or negative correlation, between fitted value and residuals, it means that residual error increase or decrease 
along with increase or decrease in predicted response variable value of the model. Another way to put it is that variance of our residual is not homogenous throughout the 
model, meaning residual value increase or decrease to match up the prediction. In the case of homoscedatic model, the variance of residual is homogenous as coefficients of 
the predictor variable accurately predict response variable value (fitted value). Thus, a horizontal trend line should be observed in ideal case where there is no 
relationship between residual and response variable and linear relationship between predictor and response variable. In our example, a clear horizontal line is observed. 

Then, normal QQ can be utilized to see if residual is normally distributed. If straight dashed line is observed, meaning that value of residual is distributed normally 
throughout the model. 

Scale location plot can be used to check homogenous variance assumption. If plot shows no distinct pattern bettern standardized residual and fitted value, which is similar 
to what is done in the first graph, then it can be assumed that homoscedacity holds true in the model. Our model gave almost straight horizontal line, hinting homoscedacity 
of the model. 


Residual vs leverage graph shows whether there are influential values that affect the model and thus must be removed. In this data, there are no points that lie outside 3 
standard deviations away and trend line is horizontal, hinting high success rate of prediction with less outliers.










***
#### (e) Check if there are outliers or influential observations 
Leverage --> Cook's distance --> DFBETAS --> Final analysis

Our final model is as below:
```{r}
final_model <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))
```

##### (1) Leverage 

= outlier?

- The larger the leverage for ith observation, the closer ith prediction is to ith outcome
- If leverage for ith observation large, then:
  - has predictor values that deviate strongly
  - possibly has large influence --> cook's distance assure this
- when p is number of unknown parameters and n is number of observations
  - Average leverage: p/n
  - Extreme leverage: larger than **2p/n**

Check for extreme leverage value:
```{r}
p <- length(coef(final_model)) ;p
n <- nrow(Data) ;n
(2*p) / n
```
- Extreme leverage when higher than 0.07035176.


```{r}
lev <- hat(model.matrix(final_model))
plot(lev, type='h')
```

```{r}
which(lev > (2*p) / n)
```
- 30 observations from the data having extreme leverage values.




***
##### (2) Influential? Cook's distance

- If Cook's distance(Di) large, then ith observation has large influence on predictions and coefficients
- Extreme Cook's distance: exceeds 50% percentile of $F_{p,n-p}$-distribution


Plot of the Cook's distance for every observation:
```{r}
plot(final_model, 4)
```


calculating 50% percentile of the F distribution
```{r}
p <- length(coef(final_model))
n <- nrow(Data)

fquantile <- qf(0.5, df1 = p, df2 = n - p) ;fquantile
```
- 50% percentile of the F distribution is 0.9097289


Investigate if cook's distance surpasses the 50% percentile of the $F_{p,n-p}$-distribution:
```{r}
cookd <- cooks.distance(final_model)
which(cookd > fquantile)
```
- According to Cook's distance, there are no observation that influences model. In other words, although there were 30 observations having high leverage, none of them are 
- influential.




***
##### (3) DFBETAS
- diagnostic measure for influence of ith observation on each regression coefficient separately
- tells where exactly or what in the model is influenced
- only calculate for points with high cook's distance 
- ```DFBETAS = (Bj-Bj(i))/SD```

Bj(i) = Bj - DFBETAS * SD(Bj)

- sign indicates if deleting observation i causes an increase(DFBETAS<0) or decrease(DFBETAS>0) in each coefficient
- Extreme DFBETAS: **exceeds 1** in small to mediocre datasets, and 2/sqrt(n) in large datasets

```{r}
2 / sqrt(n)
```
Extreme DFBETA when exceeds 0.1417762

However, we do not have any point(observation) that is influential, so there no need to calculate DFBETAS.




***
##### (4) Final analysis
There are no outliers or influential observations. It can be verified with the Residual plots.

```{r}
plot(final_model, 5)
```
```{r}
plot(final_model, 6)
```

No outlier or influential point is observed from either plots.

Hence, we now conclude that there are no outliers or influential observations.


***
Assumptions can be re-checked by residual plots
```{r}
par(mfrow=c(2, 2))
plot(final_model)
```










***
#### (f) OLS vs PGLS

OLS and PGLS have substantial similarity. However, there is one crucial difference between the two is assumption of non-independence of residual in PGLS and independence in 
OLS. PGLS takes taxonomical similiarities between related species into account for residual variance. OLS does not and it assumes that residual is independent of taxonomical 
similarities and normally distributed. In analysis of samples of related species, utilizing PGLS will be better than utilizing OLS. 










***
### 2. Results:

Regarding our final model:
```{r}
summary(final_model)
```

- Variable log(AvgBodyMass) has high p-value of 0.68347. However, as model built hierarchy and we have two related interaction terms both with p-value < 5%, 
- log(Height):log(AvgBodyMass) and log(Width):log(AvgBodyMass), we still keep the variable log(AvgBodyMass)

- Our model has multiple R-square value of 0.9938, which means about 99.39% of our observation is explained by the model. Multiple R-squared value is used for check if the 
- prediction is good enough, to avoid overfitting.


```{r}
library(jtools) # load jtools to use function summ
```

```{r}
m1 <- lm(log(ECV) ~ log(Width))
m2 <- lm(log(ECV) ~ log(Width) + log(Height))
m3 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O))
m4 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass))
m5 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass))
m6 <- lm(log(ECV) ~ log(Width) + log(Height) + log(N.O) + log(AvgBodyMass) + log(Height):log(AvgBodyMass) + log(Width):log(AvgBodyMass))

```


**Table1**
Table containing an overview of the selected regression coefficients in each of the successive models, including t value and t value.
```{r}
summ(m1, model.info = FALSE, model.fit = FALSE, digits = 6)
summ(m2, model.info = FALSE, model.fit = FALSE, digits = 6)
summ(m3, model.info = FALSE, model.fit = FALSE, digits = 6)
summ(m4, model.info = FALSE, model.fit = FALSE, digits = 6)
summ(m5, model.info = FALSE, model.fit = FALSE, digits = 6)
summ(m6, model.info = FALSE, model.fit = FALSE, digits = 6)
```

The table gives coefficients of each selected predictor variables along with their respective t value and p value. Our final model had 6 predictor variables and an 
intercept, none of them had extremely low p value. The equation derived from the model would be:


log(ECV) =  1.284933 x log(Xw) + 1.401415 x log(Xh)  + 0.307761 x log(Xn)  + -0.012753 x log(Xavg) + -0.134928 X log(Xh) x log(Xavg) + 0.122155 x log(Xw) x log (Xavg)

***
To interpret regarding **ECV**, there need **back-transformation**:

ECV =  exp(1.284933 x log(Xw) + 1.401415 x log(Xh)  + 0.307761 x log(Xn)  + -0.012753 x log(Xavg) + -0.134928 X log(Xh) x log(Xavg) + 0.122155 x log(Xw) x log (Xavg))

***
Standard error decreases over each successive addition/deletion of predictor variable, hinting higher accuracy in each addition/deletion.

***
```{r message=FALSE, warning=FALSE}
export_summs(m1, m2, m3, m4, m5, m6, digits = 4)
```




**Table 2**
```{r}
Model <- 
c('Width',
'Width + Height',
'Width + Height + N.O',
'Width + Height + N.O + AvgBodyMass',
'Width + Height + N.O + AvgBodyMass + Height:AvgBodyMass',
'Width + Height + N.O + AvgBodyMass + Height:AvgBodyMass + Width:AvgBodyMass')

R.square <- c(0.9840, 0.9915, 0.9924, 0.9925, 0.9934, 0.9938)
RSE <- c(0.2617, 0.1908, 0.1816, 0.1801, 0.1691, 0.165)


Table2 <- data.frame(Model, R.square, RSE)
Table2
```

The table 2 shows R squared value and RSE value of each successive models. Each successive addition/deletion of predictor variable led to overall increase in R squared value 
and decrease in RSE. Both hinting that model is becoming more and more successful in accurate prediction.

```{r}
plot(y = predict(final_model), x = log(ECV))
```

The figure shows linear relation between the predicted and measued ECV values. 




***
Overall analysis of the results show that model successfully predicts ECV based on 4 predictor variables and 2 interaction terms. Decrease in residual error and increase in 
R sqaured andproves that forward selection improved model's accuracy in prediction. It has showed that there prediction of the model is independent of the influence from 
categorical variable based on taxonomy, indicating that model successfullly predicts ECV of species without taking taxonomical similaritites into account. This indicates 
that utilizing OLS as done in this model instead of PGLS is actually better for the model as categorical variable holds no statistical significance in model prediction. This 
speculation may change if sample size increases for other infraclass as it was revealed that taking placenta infraclass as categorical variable revealed to be significant. 
Thus, to make sensible judgement regarding whether infraclass should be considered as categorical variable, sample size must increase. 

***


