---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
options(repos=structure(c(CRAN="https://cran.cnr.berkeley.edu/")))
```

## 1) Introduction
In 1986,a group of urologists in London published a research paper in The British Medical Journal that compared the effectiveness of two different methods to remove kidney
stones. Treatment A was open surgery (invasive), and treatment B was percutaneous nephrolithotomy (less invasive). When they looked at the results from 700 patients, 
treatment B had a higher success rate. However, when they only looked at the subgroup of patients with different kidney stone sizes, treatment A had a better success rate. 
What is going on here? This known statistical phenomenon is called Simpon’s paradox. Simpon’s paradox occurs when trends appear in subgroups but disappear or reverse when 
subgroups are combined.


## 2) Project Data Set
### 2.1 Loading the Data Set
The dataset was prepared as a text “comma-separated value” (CSV) ﬁle named KSData.csv.

***

Task 1: Start the R program and switch to the directory where the dataset is stored.

Task 2: Examine the contents of the CSV ﬁle.

```
file.show("KS_Data.csv")
```
***


#### Question 1
What is the format of the ﬁrst line? What does it represent?
```

The first line represents the header of the data, which represents the names of variables of data; treatment, stone size and success. Each represent a column of the table, 
giving three columns in total. Success is displayed as two numerical values, 0 and 1 where the former represents success and latter represents failure.

```


#### Question 2
What is the format of the following lines? What do they represent?
```
 For the following lines, the first column is in alphabet, only contained either A or B. It distinguish two treatments used in this research; A for open surgery and B for 
 Percutaneous nephrolithotomy. The second column discriminates the stone size, small or large. The last column is in number, 0 or 1, which indicates success. 0 for Failure 
 and 1 for Success. As there were 700 observations, including the first line, there are 701 lines in the data file.

```
***

Task 3: Load the readr and dplyr packages and load the dataset into R using the read.csv() method and examine its structure.

***

#### Question 3
Identify each variable from the list above. Note its data type and (if applicable) numerical precision.
```
Variable treatment is represented as string data type and gives two outputs: A and B. The second variable is stone size represented as string data type with two outputs: 
large and small. Success variable is represented as integer data type and only gives 0 and 1 as output representing, success and failure, respectively. Since, success is only 
represented as two integers (0 and 1), it is impossible and inappropriate to calculate numerical precision from given data. Therefore, it is not applicable.  

```


```{r}
# Load the readr and dplyr packages
install.packages("readr")
install.packages("dplyr")
library("readr")
library("dplyr")

# Read datasets kidney_stone_data.csv into data 
data <- read.csv(file = "KS_Data.csv")

# Take a look at the first few rows of the dataset 
head(data)
```


## 3) Problem and Objectives
### 3.1 Recreate the Treatment−Success summary table
The data contains three columns: treatment, stone_size and success. To start, we want to know which treatment had a higher success rate regardless of stone size.

***

Task 4: Create a table with the number of successes and frequency of success by each treatment using the tidyverse syntax.
```{r}
# Calculate the number and frequency of success and failure of each treatment 
data %>% 
    group_by(treatment, success) %>% 
    summarise(N = n()) %>%
    mutate(Freq = round(N/sum(N), 4))
```
***


#### Question 4
What do you observe from the treatment and success rate descriptive table? Report the success rates.
```
The table shows the number and frequency of success and failure of each treatment. According to the table, regardless of stone size, treatment A shows a success rate of 
0.7800 and treatment B shows a rate of 0.8257. Hence, treatment B seems to have a higher success rate.

```


### 3.2 Bringing stone size into the picture
Now, let’s also consider stone size and see what happens.

***

Task 5: Stratify the data into small vs. large stone subcategories and compute again the success count and rate by treatment as we did in the previous task.

The ﬁnal table will be treatment X stone size X success.
```{r}
# Calculate number and frequency of success and failure by stone size for each treatment
sum_data <-
  data %>%
  group_by(treatment, stone_size, success) %>%
  summarise(N = n()) %>%
  mutate(Freq = round(N/sum(N), 4))

# Print out the data frame we just created
print.data.frame(sum_data)
```
***


#### Question 5
What is your conclusion when the data are stratiﬁed by stone size?
```
For large kidney stones, treatment A shows a success rate of 0.7300 or 73% and treatment B shows a rate of 0.6875 or 68.75%. In large kidney stone, treatment A seems to have 
a higher success rate. For the small kidney stone, treatment A shows a success rate of 0.9310 and treatment B shows a rate of 0.8667. Even in small kidney stone, treatment A 
has a higher success rate. In conclusion, treatment B had a higher success rate when kidney stone size was not taken into account (in Question 4), but treatment A had a 
higher success rate in both sizes when kidney stone size was taken as a parameter. Furthermore, both treatments have a higher success rate in ‘small kidney stone’ than ‘large 
kidney stone’.

```


### 3.3 When in doubt, rely on a plot
Sometimes a plot is a more efﬁcient way to communicate hidden numerical information in the data.

***

Task 6: In this task, you should apply a plotting technique to reveal the hidden information.
```{r}
# Load ggplot2 
install.packages("ggplot2")
library("ggplot2")

# Create a bar plot to show stone size count within each treatment
sum_data %>% 
  ggplot(aes(x = treatment, y = N)) +
  geom_bar(aes(fill = stone_size), stat = "identity")
```


***

#### Question 6
What could you observe from the bar plot? Can we conﬁrm this hypothesis with statistical testing? If you could choose 1 statistical test to examine this, which would you choose? Argue why.
```
By observing color difference, bar plot let us easily recognize portion of ‘stone size count’ within each treatment. Each treatment was not done at the same number of each 
size of kidney stone. According to the bar plot, treatment A was performed mainly on large kidney stones. However, treatment B was performed prevalently on small kidney 
stones.

Since, overall count of large kidney stone and small kidney stone cases are almost equal to each other, but distribution of the sample to subgroup does not represent that 
homogeneity, there is a random sampling error during the experiment. The hypothesis that can be generated from such observation is that random sampling error that resulted in 
irregular distribution of samples with different stone sizes in treat A and B may have caused confounding effect and Simpson’s paradox in the result.  Thus, difference in 
trend in overall and subgroup (large stones and small stones) success rates may have rooted from possible association between success rates, treatment types and stone sizes.

This hypothesis can be confirmed using, ‘Pearson’s Chi-squared test’. Pearson’s Chi-squared test is used for analyzing the association between two categorical variables. In 
this case, they are association between treatment types and stone sizes. To be exact, it analyzes whether the risk of certain event occurring in two different groups is 
correlated. Following statements are null hypothesis and alternative hypothesis for this test.

Ho: Two categorical variables (stone size and treatment types) are independent from each other. 
HA: Two categorical variables are related to each other.

The test will analyze the resulting p-values and reject null hypothesis when p-value is less than predetermined significance level, which is 0.05. Rejection of null 
hypothesis represents that two variables under analysis are dependent to each other. 

```


### 3.4 Identify and conﬁrm the confounding variable
Now, perform the test.

Task 7: Analyze the association between stone size (i.e., case severity) and treatment assignment using a statistical test.
```{r}
# Load the broom package
install.packages("broom")
library("broom")

# Run a test
trt_ss <- chisq.test(data$treatment, data$stone_size)

# Summarize the result in tidy format
tidy(trt_ss)
```


### 3.5 Remove the confounding effect
***
#### Question 7
Are we conﬁdent that stone size/case severity is indeed the confounding variable (aka, lurking variable) in this study of kidney stone treatment and success rate? Explain why?
```
Pearson’s Chi-squared test yielded Chi-square value of 189.36 and p-value of 4.395*10^-43. The p-value was more than 10-40  fold less than predetermined significance level of 
0.05, which represents that null hypothesis is rejected and there is a correlation between the two variables under analysis. The Chi-squared value was significantly greater 
1.than 5, which is an indicator that the sample is large enough and it is appropriate to use Pearson’s Chi-squared test. The two values substantiate the argument that stone 
2.size was indeed a confounding variable in this study that caused Simpson’s paradox. 

```
***


The good news is that there are ways to get rid of the effect of the confounding variable. 

Note: Logistic regression is a mathematical process that produces results that can be interpreted as an odds ratio, and it is easy to use by any statistical package. The 
special thing about logistic regression is that it can control for numerous confounders(if there is a large enough sample size). Thus logistic regression is a mathematical 
model that can give an odds ratio which is controlled for multiple confounders. This odds ratio is known as the adjusted odds ratio, because its value has been adjusted for 
the other covariates (including confounders).

Using multiple logistic regression, remove the unwanted effect of stone size, and then tidy the output with a function from the broom package.

```{r}
# Run a multiple logistic regression
m <- glm(data=data, success ~ stone_size + treatment, family="binomial")

# Print out model coefficient table in tidy format 
tidy(m)
```

As you can see, the estimated coefﬁcients are reported, together with their standard error and a p−value.

***


#### Question 8
Since you know p−values stem from statistical tests, what exactly would be tested here? Clarify your answer using the table as an example.
```
Generalized linear model tests two predictor variables (stone size and treatment) to generate dependent variable (success) in a linear model of the form Y = aX + b. Estimate 
in the table is also called as coefficients of the predictors or ‘a’ in the function. It represents that one-unit change in coefficient generates following unit change in log 
of odds. For example, stone size estimate changes the log of odds by 1.26 per one-unit change in stone size. The p-value of this table help identify whether given estimates 
are significant. If it is a value less than predetermined value of 0.05, it is considered as significant in 95% confidence level. The table shows that the coefficients of 
small stone size significantly affect success rates relative to large stone size, which is the reference category. Positive coefficient also hints that there is a positive 
correlation between success rate and small stone size, meaning small stone size is easier to treat and cure. In the meanwhile, treatment B failed to affect the success rate 
significantly compared to its counterpart, treatment A. It represents that treatment B is not significantly better at treating and curing kidney stone in general. 
```


### 3.6 Visualize model output
We successfully ﬁt a multiple regression and pulled out the model coefﬁcient estimates! Typically (and arbitrarily), P-values below 0.05 indicate statistical signiﬁcance. 
Another way to examine whether a signiﬁcant relationship exists or not is to look at the 95% conﬁdence interval (CI) of the estimate.

***

Task 8: Create a plot to see if:
```
1. The effect of a small stone is the same as a big stone? 
2. Treatment A is as effective as treatment B?
```
***
Note: Save the model coefﬁcient table from the last task as the object tidy_m. Set up ggplot() canvas with the point estimate on the y-axis and the model terms on the x-axis. 
Calculate the upper bound of the 95% CI within geom_pointrange() and add a horizontal line at zero using geom_hline()

```{r}
# Save the tidy model output into an object 
tidy_m <- tidy(m)

# Plot the coefficient estimates with 95% CI for each term in the model
tidy_m %>% 
  ggplot(aes(x = term, y = estimate)) +
  geom_pointrange(aes(ymin = estimate - 1.96 * std.error,
                      ymax = estimate + 1.96 * std.error)) +
  geom_hline(yintercept = 0)
```


***

#### Question 9
Can you identify signiﬁcant effects using the CI? Explain how.
```
Confidence intervals were constructed using standard errors identified in the table. 1.96 was multiplied to the standard error as that represent 95% of the data in normal 
distribution. In this test, null hypothesis assumes that estimate has a normal distribution with average of zero and standard deviation of 1. Thus, if y value of 0 or 
estimate value of 0 is within the confidence interval, null hypothesis is accepted and estimate of corresponding predictor variable does not significantly affect the success 
rate compared to reference category. In this case, treatment B accepts the null hypothesis as estimate value of 0 is within the confidence level. Small stone size on the 
other hand, did not have estimate value of 0 in its confidence interval and had values ranging from approximately 0.8 to 1.74. Such results further show that it affects the 
success rate significantly compared to large stone size along with low p-value of 1.55*10^-14.

```


### 3.7 Generate insights
#### Question 10
Based on the coefﬁcient estimate plot and the model output table, there is enough information to generate insights about the study. Is treatment A superior to B after taking 
into account the effect of stone size/severity level?
```
After analyzing both coefficients estimate plot and model output table, it becomes evident that treatment B is not significantly effective in treating kidney stones compared 
to treatment A. It was evident that small stone size is easier to cure and treatment B, which had higher composition of small kidney stone samples had higher overall success 
rate only because it had easier sample to treat and cure. Thus, taking stone size into account, treatment B is not a better treatment method compared to treatment A and 
higher overall success rate is only attributed by higher composition of small kidney stone samples. 

```
***


Task9: Everything is in the output table from the regression model. Recall,a coefﬁcient represents the effect size of the speciﬁc model term. A positive coefﬁcient means that 
the term is positively related to the outcome. For categorical predictors, the coefﬁcient is the effect on the outcome relative to the reference category. In our case, stone 
size large and treatment A are the reference categories.

```{r}
# Is small stone more likely to be a success after controlling for treatment option effect?
small_high_success <- "Yes"

# Is treatment A significantly better than B?
A_B_sig <- "No"
```

